---
title: "Lab 05 - La Quinta is Spanish for next to Denny's, Pt. 2"
author: "Thomas Huang"
date: "2026-02-06"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox) 
```

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
data(dennys, package = "dsbox")
data(laquinta, package = "dsbox")
```

### Exercise 1

There are three Denny's and two La Quinta in Alaska.
```{r Ex1}
dn_ak <- dennys %>%
  filter(state == "AK")
nrow(dn_ak)

lq_ak <- laquinta %>%
  filter(state == "AK")
nrow(lq_ak)
```

### Exercise 2

There are three Denny's and two La Quinta in Alaska. Each Denny's corresponds to two La Quinta. Therefore, there should be $$3\times2=6$$ pairs.

### Exercise 3

```{r Ex3}
dn_lq_ak <- full_join(dn_ak, lq_ak,
  by = "state"
)
dn_lq_ak
```

### Exercise 4

There are six observations in this data frame. The variable are street address, city, state, zip code, longitude, and latitude of each Denny's and La Quinta.
```{r Ex4}
glimpse(dn_lq_ak)
```

### Exercise 5

dplyr::mutate(new_var = ...) can be used to create a new variable.
```{r Ex5}
# Define a cool function that calculates the distance between any two points on the earth.
haversine <- function(long1, lat1, long2, lat2, round = 3) {
  # convert to radians
  long1 <- long1 * pi / 180
  lat1 <- lat1 * pi / 180
  long2 <- long2 * pi / 180
  lat2 <- lat2 * pi / 180

  R <- 6371 # Earth mean radius in km

  a <- sin((lat2 - lat1) / 2)^2 + cos(lat1) * cos(lat2) * sin((long2 - long1) / 2)^2
  d <- R * 2 * asin(sqrt(a))

  return(round(d, round)) # distance in km
}
```

### Exercise 6

```{r Ex6}
dn_lq_ak <- dn_lq_ak %>% 
  dplyr::mutate(
    distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y)
  )

glimpse(dn_lq_ak)
```

Add exercise headings as needed.
